---
title: "African Logistics Performance Optimization"
author: "Franck Djandja"
output: html_document
---

# Project Overview

This project analyzes logistics performance across several African countries to identify delivery delays, transportation inefficiencies, and opportunities for operational optimization.

# Libraries

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
```

# Load Dataset

```{r}
data <- read.csv("../data/logistics_dataset.csv")
head(data)
```

# Data Preparation

```{r}
data$Delay_days <- data$Delivery_Time_days - data$Expected_Delivery_days
data$Cost_per_km <- data$Shipping_Cost / data$Distance_km
```

# Shipments by Country

```{r}
data %>%
  count(Country) %>%
  ggplot(aes(x=reorder(Country,n), y=n)) +
  geom_bar(stat="identity") +
  coord_flip() +
  labs(title="Total Shipments by Country",
       x="Country",
       y="Number of Shipments")
```

# Delivery Delay by Country

```{r}
data %>%
  group_by(Country) %>%
  summarise(avg_delay = mean(Delay_days)) %>%
  ggplot(aes(x=reorder(Country,avg_delay), y=avg_delay)) +
  geom_bar(stat="identity") +
  coord_flip() +
  labs(title="Average Delivery Delay by Country",
       x="Country",
       y="Average Delay (days)")
```

# Delivery Delay by Transport Mode

```{r}
data %>%
  group_by(Transport_Mode) %>%
  summarise(avg_delay = mean(Delay_days)) %>%
  ggplot(aes(x=Transport_Mode, y=avg_delay)) +
  geom_bar(stat="identity") +
  labs(title="Delivery Delay by Transport Mode",
       x="Transport Mode",
       y="Average Delay (days)")
```

# Cost Efficiency

```{r}
ggplot(data, aes(x=Distance_km, y=Shipping_Cost)) +
  geom_point(alpha=0.3) +
  labs(title="Shipping Cost vs Distance",
       x="Distance (km)",
       y="Shipping Cost")
```

# Shipment Trend

```{r}
data$Order_Date <- as.Date(data$Order_Date)

shipments_ts <- data %>%
  count(Order_Date)

ggplot(shipments_ts, aes(x=Order_Date, y=n)) +
  geom_line() +
  labs(title="Shipment Volume Over Time",
       x="Date",
       y="Number of Shipments")
```

# Key Insights

- Road transport accounts for the majority of shipments.
- Certain countries experience higher delivery delays.
- Shipping costs increase with distance but reveal operational inefficiencies.
- Shipment demand shows a gradual increase over time.
